---
title: 'Code Notebook for : </br>  </br> Developmental Emergence of Neurogliaform
  Cell Diversity'
author: "Lucia Gomez Teijeiro"
date: "2023-07-19"
output:
  html_document:
    df_print: paged
subtitle: "</br>  </br> Embryonic Pseudotime - CGE & POA </br>  </br>"
---

 </br>
**Information for replication:**  </br>
This code was executed using Seurat version 3 and R version 3.6.1. 

```{=html}
<style>
p {
    font-size: 15px;
    line-height: 20px;
    margin: 0px 0px 12px 0px;
}
h1, h2, h3, h4, h5, h6, legend {
    font-family: Montserrat, sans-serif;
    font-weight: 700;
    font-size: 20px;
    color:#56088a;
}
body {
  font-family: "Montserrat", serif, "Open Sans", sans-serif;
  max-width:5000px !important;
}


#sidebar .date {
  color: white;
  font-size:70%;
}
#sidebar h2 {
    z-index: 200;
    background-color: #56088a;
    text-align: center;
    padding: 0.7em;
    display: block;
    color: #fcfcfc;
    font-size: 100%;
    margin-top: 0px;
    margin-bottom: 0.809em;
}
#sidebar h2 a {
  font-family:"Montserrat", serif, "Open Sans", sans-serif;
  color: white;
  font-weight:700;
  font-size:120%;
  padding: 0.1em;
  line-height: 25px;
  margin-top: 20px;
  margin-bottom: 20px;
  margin-left:5px;
  margin-left:5px;
}
#nav-top span.glyphicon {
    color: #56088a;
    font-size: 28px;
}
#content a{
  color:white;
  font-size:110%;
  font-weight:700;
  margin-top: 30px;
}
#postamble {
  background:#56088a;
  border-top:solid 3px #56088a;
  width:100%;
}
</style>
```
```{css, echo=FALSE}
.title {
  text-align: center;
}
.subtitle {
  text-align: center;
}
@media screen and (max-width: 5000px){
#content{
    max-width:5000px;
    margin-left: 300;
    margin-right:0;
    padding: 1em;
    padding-top: 1em;
    padding-right: 8em;
    padding-bottom: 1em;
    padding-left: 25em;
    background:white;
}}
#sidebar{ 
  background: black;
  width:32%;
}

```


```{r setup, include=TRUE, eval=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load required packages
library(rmdformats) ; library(Biobase);library(Seurat); library(SummarizedExperiment)
library(princurve) ; library(scater) ; library(ggplot2) ; library(reshape2) ; library(Biobase)
```


## Load data 


```{r load, include=TRUE, eval=FALSE}

# load embryonic datasets after QC 

POA_DP_E14 <- readRDS("~/Desktop/PHD/OSF/data/POA_DP_E14.rds")
POA_FT_E14 <- readRDS("~/Desktop/PHD/OSF/data/POA_FT_E14.rds")
POACGE_10X_E14 <- readRDS("~/Desktop/PHD/OSF/data/POACGE_10X_E14.rds")

```


## Find Variable Genes across all E14 datasets excluding variability derived from eminence, aiming at maturation genes 


```{r vargenes, include=TRUE, eval=FALSE}

# calculate vargenes in CGE10X and POA10X separatedly and intersect the lists
    # this will remove genes that are specific for POA or CGE (since we only want genes that explain a common maturation trajectory)
# calculate variable genes in CCHT9 and CCHT10 separatedly and paste the vectors - since CCHT9 is prog and CCHT10 is prec
    # keep only the genes that are intersection between 10x intersected and the union of the chips
features <- local({
  vargenes <- function(object){
    object <- NormalizeData(object)
    features <- FindVariableFeatures(object,selection.method="mvp")
    features <- features@assays$RNA@var.features
  }
  vargenesCGE10X <- vargenes(subset(POACGE_10X_E14,experiment=="CGE"))
  vargenesPOA10X <- vargenes(subset(POACGE_10X_E14,experiment=="POA"))
  vargenes10X <- intersect(vargenesCGE10X,vargenesPOA10X)
  vargenesCCHT9 <- vargenes(POA_DP_E14)
  vargenesCCHT10 <- vargenes(POA_FT_E14)
  vargenesCHIPS <- c(vargenesCCHT9,vargenesCCHT10)
  features <- intersect(vargenes10X,vargenesCHIPS)
})
# saveRDS(features,"out/E14_features_pseudotime.rds")
features <- readRDS("out/E14_features_pseudotime.rds")

```


## Merge all E14 data and calculate cell cycle score for each cell


```{r cellcycle, include=TRUE, eval=FALSE}

E14_seurat <- merge(POACGE_10X_E14,list(POA_DP_E14,POA_FT_E14))
cellcycle <- function(data){
  cell_cycle_genes_s <- readRDS("data/cell_cycle_genes_s.rds")
  cell_cycle_genes_g2m <- readRDS("data/cell_cycle_genes_g2m.rds")
  data <- NormalizeData(data)
  data <- CellCycleScoring(data,s.features = cell_cycle_genes_s, g2m.features = cell_cycle_genes_g2m)
}
E14_seurat <- cellcycle(E14_seurat)

```


## Calculate Pseudotime Score for each cell (position in pca trajectory defined by maturation genes)


```{r pseudotime, include=TRUE, eval=FALSE}

# assemble data into expression set object
make.eset <- function() {
  data <- E14_seurat
  sce <- ExpressionSet(as.matrix(data@assays$RNA@data),phenoData = AnnotatedDataFrame(data@meta.data))
  assayDataElement(sce,"scale")<-as.matrix(data@assays$RNA@data)
  sce
}
sce <- make.eset()

# Define function for calculate pca projection on maturation genes and retrieve components explaining at least 3% of the variance of maturation
pca_map <- function(m,features) {
  pca <- prcomp(m[,features])
  eigen <- pca$sdev^2
  pca$x <- pca$x[,which(eigen[1:50]/sum(eigen[1:50])>0.03)]
  pca$x
}

# define function for calculating pseudotime - pca, principal curve fitting, mobile mean and progenitor - postmitotic boundary
my_processing_pseudotime <- function(sce,features=features,genes_begin="Nes") {
  data <- t(assayDataElement(sce,"scale"))
  sce$pca <- pca_map(data,features)
  ### fit a principal curve on significant PC components
  fit1 <- principal_curve(as.matrix(sce$pca),smoother="lowess")
  sce$fit1.s <- fit1$s
  sce$pseudotime <- as.numeric(fit1$lambda)
  sce$pseudotime <- (sce$pseudotime-min(sce$pseudotime))/max(sce$pseudotime-min(sce$pseudotime))
  begin <- exprs(sce[,order(sce$pseudotime)])[genes_begin,]
  sce$pseudotime <- ifelse(rep(which(c(mean(begin[1:5]),mean(begin[length(begin)-5:length(begin)]))==pmax(mean(begin[1:5]),mean(begin[length(begin)-5:length(begin)])))==1,length(sce$pseudotime)),sce$pseudotime,abs(sce$pseudotime-1))
  
  # mobile mean
  sce_order <- sce[,order(sce$pseudotime)]
  sce_order$mm <- caTools::runmean(sce_order$Phase%in%c("S","G2M"),1000)
  
  # determine progenitor / post-mitotic boundary in pseudotime
  i <- which(sce_order$mm < mean(sce_order$mm)/2)[1]
  sce_order$mitotic <- ifelse(sce_order$pseudotime<sce_order$pseudotime[i],"Prog","Post")
  sce_order
}

sce <- my_processing_pseudotime(sce,features=features,genes_begin="Nes")

```





```{r save, include=TRUE, eval=FALSE}

# Add pseudotime calculations to seurat object and SAVE
local({
  sce <- sce
  data <- E14_seurat
  data@meta.data$mitotic <- data@meta.data$mm <- data@meta.data$pseudotime <- NA
  i <- match(colnames(sce),rownames(data@meta.data))
  data@meta.data$pseudotime[i] <- sce$pseudotime
  data@meta.data$mm[i] <- sce$mm
  data@meta.data$mitotic[i] <- sce$mitotic
  #saveRDS(data,"out/E14_pseudotime.rds")
})


```




