---
title: 'Code Notebook for : </br>  </br> Developmental Emergence of Neurogliaform
  Cell Diversity'
author: "Lucia Gomez Teijeiro"
date: "2023-07-19"
output:
  html_document:
    df_print: paged
subtitle: "</br>  </br> Embryonic NGC identification (time-conserved) and NGC conserved genes </br>  </br>"
---

 </br>
**Information for replication:**  </br>
This code was executed using Seurat version 3 and R version 3.6.1. 

```{=html}
<style>
p {
    font-size: 15px;
    line-height: 20px;
    margin: 0px 0px 12px 0px;
}
h1, h2, h3, h4, h5, h6, legend {
    font-family: Montserrat, sans-serif;
    font-weight: 700;
    font-size: 20px;
    color:#56088a;
}
body {
  font-family: "Montserrat", serif, "Open Sans", sans-serif;
  max-width:5000px !important;
}


#sidebar .date {
  color: white;
  font-size:70%;
}
#sidebar h2 {
    z-index: 200;
    background-color: #56088a;
    text-align: center;
    padding: 0.7em;
    display: block;
    color: #fcfcfc;
    font-size: 100%;
    margin-top: 0px;
    margin-bottom: 0.809em;
}
#sidebar h2 a {
  font-family:"Montserrat", serif, "Open Sans", sans-serif;
  color: white;
  font-weight:700;
  font-size:120%;
  padding: 0.1em;
  line-height: 25px;
  margin-top: 20px;
  margin-bottom: 20px;
  margin-left:5px;
  margin-left:5px;
}
#nav-top span.glyphicon {
    color: #56088a;
    font-size: 28px;
}
#content a{
  color:white;
  font-size:110%;
  font-weight:700;
  margin-top: 30px;
}
#postamble {
  background:#56088a;
  border-top:solid 3px #56088a;
  width:100%;
}
</style>
```
```{css, echo=FALSE}
.title {
  text-align: center;
}
.subtitle {
  text-align: center;
}
@media screen and (max-width: 5000px){
#content{
    max-width:5000px;
    margin-left: 300;
    margin-right:0;
    padding: 1em;
    padding-top: 1em;
    padding-right: 8em;
    padding-bottom: 1em;
    padding-left: 25em;
    background:white;
}}
#sidebar{ 
  background: black;
  width:32%;
}

```


```{r setup, include=TRUE, eval=FALSE}

knitr::opts_chunk$set(echo = TRUE)
# Load required packages
library(Seurat) ; library(URD) ; library(readr) ; library(venn) ; library(ggplot2) ; library(readr) ; library(bmrm)

```


## Load POA integration data, postnatal type architecture genes (NGC vs nonNGC) and postnatal merged data


```{r load, include=TRUE, eval=FALSE}

E14_POA_integration <- readRDS("out/E14_POA_integration.rds")
NGCvsnonNGC <- read.csv("out/NGCvsnonNGC_acrossdev_fullrank.csv")
P15P30P56_f <- readRDS("out/P15P30P56_f.rds") ; P15P30P56_f <- UpdateSeuratObject(P15P30P56_f) # Update object to v3

```



## Calculate NGC pseudogene - mean expression for all genes conserved between postnatal NGCs and embryonic POA cells


```{r pseudogene, include=TRUE, eval=FALSE}

# calculate mean expression for each gene expressed in POA data (E14)
POAgenemeans <- rowMeans(as.matrix(E14_POA_integration@assays$RNA@data)) 
POAgenemeans <- POAgenemeans[POAgenemeans>0.05]

# find top 150 postnatal NGC genes (SVM weights)
NGC_svm <- NGCvsnonNGC[NGCvsnonNGC$w>0,]
NGC_svm <- NGC_svm[order(NGC_svm$w,decreasing=TRUE),]
NGC_svm <- NGC_svm[c(1:150),]

# find intersection between postnatal and embryonic gene lists - constitutive NGC genes
setlist <- venn(list(adult=NGC_svm$X,embry=names(POAgenemeans)))
setlist_int <- attr(setlist,"intersections")
NGCgenes <-  c(setlist_int$`adult:embry`)

# calculate NGC pseudogene - normalize by total genes expressed per cell
E14_POA_integration$NGCpseudogene <- (colMeans(as.matrix(E14_POA_integration@assays$RNA@data[NGCgenes,])) / E14_POA_integration$nFeature_RNA ) * 10000

```


## Indentify NGC cells in POA data (above 80 percentile in NGC pseudogene and belonging to cluster 2 - Hmx3 expressing or any progenitor cluster)


```{r ngcs, include=TRUE, eval=FALSE}


# IDENTIFY NGCs 
# calculate NGC pseudogene factor - bigger than the 80 percentile and in the path leading to blue cluster
E14_POA_integration$NGCpseudogene_f <- ifelse(E14_POA_integration$NGCpseudogene >= quantile(E14_POA_integration$NGCpseudogene[E14_POA_integration$clusts_ordered %in% c(0,1,2,5)],0.80,na.rm=TRUE) & 
                                   E14_POA_integration$clusts_ordered %in% c(0,1,2,5),"NGC","other") 

```


## Find time-conserved NGC genes (from E14 to P56) - Z-score 


```{r mu, include=TRUE, eval=FALSE}

mu <- t(rowmean(t(as.matrix(E14_POA_integration@assays$RNA@data[NGCgenes,])),E14_POA_integration$NGCpseudogene_f))
E14mu <- mu[,"NGC"] - mu[,"other"] 

mu2 <- t(rowmean(t(as.matrix(P15P30P56_f@assays$RNA@data[NGCgenes,])),P15P30P56_f$NGC))
PNmu <- mu2[,"NGC"] - mu2[,"nonNGC"] ;

# normalize by sd and assemble in DF
E14mu <- E14mu/sd(E14mu)
PNmu <- PNmu/sd(PNmu)
mudf <- data.frame(E14mu=E14mu,PNmu=PNmu)

# save data with Zscore for NGC genes
#write.csv(mudf,"out/NGCgenes_zscore_E14P15P30P56.csv")

```






