---
title: 'Code Notebook for : </br>  </br> Developmental Emergence of Neurogliaform Cell Diversity'
author: "Lucia Gomez Teijeiro"
date: "2023-07-19"
output:
  html_document:
    df_print: paged
subtitle: </br>  </br> Postnatal Molecular Architecture - NGC Subtypes (Dock5 vs Lsp1) </br>  </br> 
---

 </br>
**Information for replication:**  </br>
This code was executed using Seurat version 2 and R version 3.4.2. 
Some of the functions here used are no longer supported in current versions.

```{=html}
<style>
p {
    font-size: 15px;
    line-height: 20px;
    margin: 0px 0px 12px 0px;
}
h1, h2, h3, h4, h5, h6, legend {
    font-family: Montserrat, sans-serif;
    font-weight: 700;
    font-size: 20px;
    color:#56088a;
}
body {
  font-family: "Montserrat", serif, "Open Sans", sans-serif;
  max-width:5000px !important;
}


#sidebar .date {
  color: white;
  font-size:70%;
}
#sidebar h2 {
    z-index: 200;
    background-color: #56088a;
    text-align: center;
    padding: 0.7em;
    display: block;
    color: #fcfcfc;
    font-size: 100%;
    margin-top: 0px;
    margin-bottom: 0.809em;
}
#sidebar h2 a {
  font-family:"Montserrat", serif, "Open Sans", sans-serif;
  color: white;
  font-weight:700;
  font-size:120%;
  padding: 0.1em;
  line-height: 25px;
  margin-top: 20px;
  margin-bottom: 20px;
  margin-left:5px;
  margin-left:5px;
}
#nav-top span.glyphicon {
    color: #56088a;
    font-size: 28px;
}
#content a{
  color:white;
  font-size:110%;
  font-weight:700;
  margin-top: 30px;
}
#postamble {
  background:#56088a;
  border-top:solid 3px #56088a;
  width:100%;
}
</style>
```
```{css, echo=FALSE}
.title {
  text-align: center;
}
.subtitle {
  text-align: center;
}
@media screen and (max-width: 5000px){
#content{
    max-width:5000px;
    margin-left: 300;
    margin-right:0;
    padding: 1em;
    padding-top: 1em;
    padding-right: 8em;
    padding-bottom: 1em;
    padding-left: 25em;
    background:white;
}}
#sidebar{ 
  background: black;
  width:32%;
}

```


```{r setup, include=TRUE, eval=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load required packages
library(rmdformats) ; library(Seurat) ; library(bmrm) ; library(parallel) ; library(readr) ; library(abind)
```


## Load data and subset to keep just NGC cells

```{r load, include=TRUE, eval=FALSE}

P15P30P56_f <- readRDS("out/P15P30P56_f.rds")
P15P30P56_f_DL <- SubsetData(P15P30P56_f,cells.use = P15P30P56_f@meta.data$identity %in% c("Lamp5 Plch2 Dock5","Lamp5 Lsp1"))

```


## Binary classification by cell SUBTYPE to find genes characterizing types (NGC vs nonNGC)

```{r classif, include=TRUE, eval=FALSE}

# Scale data regressing out expression effects due age variations (we want to discover the genes with stable expression over time for each cell type)
P15P30P56_f_DL <- ScaleData(P15P30P56_f_DL,vars.to.regress = c("nGene","nUMI","protocol"))
# define boolean vector for type (target of the binary classification)
Dock5vsLsp1 <-  ifelse(P15P30P56_f_DL@meta.data$identity=="Lamp5 Plch2 Dock5",TRUE,FALSE)

# load machine learning library - containing code for the hingeloss SVM classifier
source("src/lib/ML_lib.R")
# run classification
X <- t(as.matrix(P15P30P56_f_DL@scale.data)) 
folds <- balanced.cv.fold(Dock5vsLsp1,num.cv=10) ; table(folds,Dock5vsLsp1)
Dock5vsLsp1_hinge <- mclapply(levels(folds),mc.cores=1,function(f) {
  Y <- feature.selected.hinge(x.trn=X[folds!=f,],y.trn=Dock5vsLsp1[folds!=f],x.tst=X,LAMBDA=2,LAMBDA2=2,n=1000,MAX_ITER=1000)
  D <- attr(Y,"decision.value")
  D[folds!=f] <- NA
  R <- attr(Y,"rank")
  rownames(R) <- colnames(X)
  return(list("decision"=D,"rank"=R))
})

# Split list to create a matrix with decisions (each CV is one column) and a list with ranks (each CV is a list element)
Dock5vsLsp1C_hinge_decisionmatrix <- sapply(Dock5vsLsp1_hinge,"[[",1)
Dock5vsLsp1_ranklist <- lapply(Dock5vsLsp1_hinge,"[[",2)
# Calculate global decision value
Dock5vsLsp1_hinge_globaldecision <- rowSums(Dock5vsLsp1C_hinge_decisionmatrix,na.rm = T)

# Calculate global rank per gene - robust genes for each type because they are commonly predicted in the 5 CVs
# First I transform rank list into a 3D array so each rank matrix (one per CV) is in a different 3d level of the array
Dock5vsLsp1_hinge_rankarray <- abind(Dock5vsLsp1_ranklist,along=3)
# I want to calculate the mean value for all columns in rank dataframe - p value
# Calculate significancy of our model to a random model using the same number of cells, for cell vectors with NGCvsNonNGC_ranklist logical value 
Dock5vsLsp1_robustrank_commontoCVs <- rowSums(Dock5vsLsp1_hinge_rankarray,dims = 2) 
Dock5vsLsp1_robustrank_commontoCVs_df <- as.data.frame(Dock5vsLsp1_robustrank_commontoCVs)
Dock5vsLsp1_robustrank_commontoCVs_df_div10 <- Dock5vsLsp1_robustrank_commontoCVs_df/10

# save full rank object (only 150 top and 150 bottom rows must be used for DEG functional enrichment analysis - GO, HGNC)
#write.csv(Dock5vsLsp1_robustrank_commontoCVs_df_div10,file="out/Dock5vsLsp1_acrossdev_fullrank.csv")

```


## For calculating GO and HGNC enrichment analysis use the lib "lib/go_hgnc.R" 


