---
title: 'Code Notebook for : </br>  </br> Developmental Emergence of Neurogliaform Cell Diversity'
author: "Lucia Gomez Teijeiro"
date: "2023-07-19"
output:
  html_document:
    df_print: paged
subtitle: </br>  </br> Postnatal Molecular Architecture - NGC Type </br>  </br> 
---

 </br>
**Information for replication:**  </br>
This code was executed using Seurat version 2 and R version 3.4.2. 
Some of the functions here used are no longer supported in current versions.

```{=html}
<style>
p {
    font-size: 15px;
    line-height: 20px;
    margin: 0px 0px 12px 0px;
}
h1, h2, h3, h4, h5, h6, legend {
    font-family: Montserrat, sans-serif;
    font-weight: 700;
    font-size: 20px;
    color:#56088a;
}
body {
  font-family: "Montserrat", serif, "Open Sans", sans-serif;
  max-width:5000px !important;
}


#sidebar .date {
  color: white;
  font-size:70%;
}
#sidebar h2 {
    z-index: 200;
    background-color: #56088a;
    text-align: center;
    padding: 0.7em;
    display: block;
    color: #fcfcfc;
    font-size: 100%;
    margin-top: 0px;
    margin-bottom: 0.809em;
}
#sidebar h2 a {
  font-family:"Montserrat", serif, "Open Sans", sans-serif;
  color: white;
  font-weight:700;
  font-size:120%;
  padding: 0.1em;
  line-height: 25px;
  margin-top: 20px;
  margin-bottom: 20px;
  margin-left:5px;
  margin-left:5px;
}
#nav-top span.glyphicon {
    color: #56088a;
    font-size: 28px;
}
#content a{
  color:white;
  font-size:110%;
  font-weight:700;
  margin-top: 30px;
}
#postamble {
  background:#56088a;
  border-top:solid 3px #56088a;
  width:100%;
}
</style>
```
```{css, echo=FALSE}
.title {
  text-align: center;
}
.subtitle {
  text-align: center;
}
@media screen and (max-width: 5000px){
#content{
    max-width:5000px;
    margin-left: 300;
    margin-right:0;
    padding: 1em;
    padding-top: 1em;
    padding-right: 8em;
    padding-bottom: 1em;
    padding-left: 25em;
    background:white;
}}
#sidebar{ 
  background: black;
  width:32%;
}

```


```{r setup, include=TRUE, eval=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load required packages
library(rmdformats) ; library(Seurat) ; library(bmrm) ; library(parallel) ; library(readr) ; library(abind)
```


## Load data

```{r load, include=TRUE, eval=FALSE}

Tasic18 <- readRDS("data/Tasic18_Seurat.rds")
P15 <- readRDS("data/NGC_P15_Seurat.rds")
P30 <- readRDS("data/NGC_P30_Seurat.rds")
ID_P15 <- read.csv("out/CCA_KNN_SVM_T18_P15.csv")
ID_P30 <- read.csv("out/CCA_KNN_SVM_T18_P30.csv")

```


## Subset & Merge data to keep a random sample balanced per cell type (NGC vs nonNGC) (apply QC on fluorescence)


```{r sample, include=TRUE, eval=FALSE}

# Annotate datasets by cell subtype
P15@meta.data$identity <- ID_P15$IDconsensus[match(colnames(P15),ID_P15$cellnames)]
P30@meta.data$identity <- ID_P30$IDconsensus[match(colnames(P30),ID_P30$cellnames)]
Tasic18@meta.data$identity <- Tasic18@meta.data$cluster

# Keep only GFP+TOM+ in Dock5vsLsp1 subtypes and GFP+TOM- in the rest of the clusters
  # remove TOM- cells from NGC clusters
P15_1 <- SubsetData(P15, cells.use = (P30@meta.data$Tom_cells %in% FALSE & P30@meta.data$identity %in% c("Lamp5 Lsp1","Lamp5 Plch2 Dock5")))
P30_1 <- SubsetData(P30, cells.use = (P30@meta.data$Tom_cells %in% FALSE & P30@meta.data$identity %in% c("Lamp5 Lsp1","Lamp5 Plch2 Dock5")))
  # remove TOM+ cells from nonNGCclusters
P15_2 <- SubsetData(P15, cells.use = (P15@meta.data$Tom_cells %in% TRUE & !(P15@meta.data$identity %in% c("Lamp5 Lsp1","Lamp5 Plch2 Dock5"))))
P30_2 <- SubsetData(P30, cells.use = (P30@meta.data$Tom_cells %in% TRUE & !(P30@meta.data$identity %in% c("Lamp5 Lsp1","Lamp5 Plch2 Dock5"))))
  # ReMerge datasets
P15 <- MergeSeurat(P15_1,P15_2) ; P30 <- MergeSeurat(P30_1,P30_2)

# Calculate how many cells per subtype must be used from Tasic dataset in order to balance classes (subtypes)
  # number of cells to select per subtype will be determined by the mean cell per subtype between P15 and P30 datasets
    # subtypes with less than 5 cells will be discarded
P15P30P56 <- MergeSeurat(P15,P30,min.cells = 5) ; P15P30P56 <- MergeSeurat(P15P30P56,Tasic18,min.cells = 5)
set.seed(123)
n <- round(rowMeans(table(P15P30P56@meta.data$identity,P15P30P56@meta.data$protocol)[,c("NkxP15","NkxP30")]),digits = 0)
n <- pmax(n,1)
cells.to.keep <- unlist(mapply(function(lev,num){sample(Tasic18@cell.names[Tasic18@meta.data$identity %in% lev],num)},names(n),n))
P15P30P56_f <- SubsetData(P15P30P56,cells.use = (P15P30P56@cell.names %in% cells.to.keep) | (P15P30P56@meta.data$protocol %in% c("NkxP15","NkxP30")))

# Create a factor with ngc (dock5+lsp1) and nonNGC (other subtypes)
P15P30P56_f@meta.data$NGC <- as.factor(P15P30P56_f@meta.data$identity)

```


## Reconstruct maturation order for cells (ordinal machine learning model)


```{r ordi, include=TRUE, eval=FALSE}

x <- as.matrix(t(P15P30P56_f@scale.data))
x <- scale(x,center=TRUE,scale=FALSE)

P15P30P56_f@meta.data$age.ordi <- as.factor(P15P30P56_f@meta.data$protocol)
levels(P15P30P56_f@meta.data$age.ordi) <- c("1","2","3")
P15P30P56_f@meta.data$age.ordi <- as.integer(as.character(P15P30P56_f@meta.data$age.ordi))
table(P15P30P56_f@meta.data$age.ordi,P15P30P56_f@meta.data$protocol)

set.seed(1234)
folds <- balanced.cv.fold(P15P30P56_f@meta.data$age.ordi,num.cv=10)
y <- P15P30P56_f@meta.data$age.ordi
ordipred.cv.cells <- simplify2array(mclapply(levels(folds),mc.cores=1,function(f) {
  w <- nrbm(ordinalRegressionLoss(x[folds!=f,],P15P30P56_f@meta.data$age.ordi[folds!=f]),LAMBDA = 105,EPSILON_TOL = 1e-7,MAX_ITER = 100)
  Y <- predict(w,x)
  Y[folds!=f] <- NA
  Y
}))
P15P30P56_f@meta.data$ordi.cellscore.age.cv <- rowSums(ordipred.cv.cells,na.rm=TRUE)
P15P30P56_f@meta.data$ordi.cellscore.age.cv <- -(P15P30P56_f@meta.data$ordi.cellscore.age.cv)

# Waves plot (expression of a gene through maturation)
df <- as.data.frame(as.matrix(t(P15P30P56_f@data)))
df$type <- as.factor(P15P30P56_f@meta.data$NGC)
df$cellscoreage.cv <- as.numeric(P15P30P56_f@meta.data$ordi.cellscore.age.cv)
ggplot(df)+ scale_color_manual(values=c("green", "yellow")) + scale_fill_manual(values=c("green", "yellow")) +
  geom_point(aes(df$cellscoreage.cv,df$Tox2,fill=as.factor(df$type)),shape=21,colour="black",alpha=1,size=2.5) + 
  geom_smooth(aes(df$cellscoreage.cv,df$Tox2,color=df$type),method="loess", span=1.5)

# Save object with 3 posnatal ages, cell selection ordipred and ngc vector
#saveRDS(P15P30P56_f,file="P15P30P56_f.rds")

```


## Binary classification by cell TYPE to find genes characterizing types (NGC vs nonNGC)

```{r classif, include=TRUE, eval=FALSE}

# Scale data regressing out expression effects due age variations (we want to discover the genes with stable expression over time for each cell type)
P15P30P56_f <- ScaleData(P15P30P56_f,vars.to.regress = c("nGene","nUMI","protocol"))
# define boolean vector for type (target of the binary classification)
NGCvsNonNGC <-  ifelse(P15P30P56_f@meta.data$NGC=="NGC",TRUE,FALSE)

# load machine learning library - containing code for the hingeloss SVM classifier
source("src/lib/ML_lib.R")
# run classification
X <- t(as.matrix(P15P30P56_f@scale.data)) 
folds <- balanced.cv.fold(NGCvsNonNGC,num.cv=10) ; table(folds,NGCvsNonNGC)
NGCvsNonNGC_hinge <- mclapply(levels(folds),mc.cores=1,function(f) {
  Y <- feature.selected.hinge(x.trn=X[folds!=f,],y.trn=NGCvsNonNGC[folds!=f],x.tst=X,LAMBDA=2,LAMBDA2=2,n=1000,MAX_ITER=1000)
  D <- attr(Y,"decision.value")
  D[folds!=f] <- NA
  R <- attr(Y,"rank")
  rownames(R) <- colnames(X)
  return(list("decision"=D,"rank"=R))
})

# Split list to create a matrix with decisions (each CV is one column) and a list with ranks (each CV is a list element)
NGCvsNonNGC_hinge_decisionmatrix <- sapply(NGCvsNonNGC_hinge,"[[",1)
NGCvsNonNGC_ranklist <- lapply(NGCvsNonNGC_hinge,"[[",2)
# Calculate global decision value
NGCvsNonNGC_hinge_globaldecision <- rowSums(NGCvsNonNGC_hinge_decisionmatrix,na.rm = T)

# Calculate global rank per gene - robust genes for each type because they are commonly predicted in the 5 CVs
# First I transform rank list into a 3D array so each rank matrix (one per CV) is in a different 3d level of the array
NGCvsNonNGC_hinge_rankarray <- abind(NGCvsNonNGC_ranklist,along=3)
# I want to calculate the mean value for all columns in rank dataframe - p value
# Calculate significancy of our model to a random model using the same number of cells, for cell vectors with NGCvsNonNGC_ranklist logical value 
NGCvsNonNGC_robustrank_commontoCVs <- rowSums(NGCvsNonNGC_hinge_rankarray,dims = 2) 
NGCvsNonNGC_robustrank_commontoCVs_df <- as.data.frame(NGCvsNonNGC_robustrank_commontoCVs)
NGCvsNonNGC_robustrank_commontoCVs_df_div10 <- NGCvsNonNGC_robustrank_commontoCVs_df/10

# save full rank object (only 150 top and 150 bottom rows must be used for DEG functional enrichment analysis - GO, HGNC)
#write.csv(NGCvsNonNGC_robustrank_commontoCVs_df_div10,file="out/NGCvsnonNGC_acrossdev_fullrank.csv")

# save universe of genes expressed in P15P30P56_f object (universe for DEG functional analysis)
#write(rownames(P15P30P56_f@data),file="universe.txt")

```


## For calculating GO and HGNC enrichment analysis use the lib "lib/go_hgnc.R" 





